### 2016-01-03 article/article.tex
New Algorithm ! Ongoing test and performance improvement

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={SODA POP: SOft solving and Defect Aware Partial Ordering Planning},
            pdfauthor={Paper ID\#232},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{listings}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

\title{SODA POP: SOft solving and Defect Aware Partial Ordering Planning}
\author{Paper ID\#232}
\date{}
\usepackage{xcolor, listings}
\usepackage{latex/aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}

% We now define the sixteen \solarized{} colors.
\definecolor{solarized-base03} {RGB}{000, 043, 054}
\definecolor{solarized-base02} {RGB}{007, 054, 066}
\definecolor{solarized-base01} {RGB}{088, 110, 117}
\definecolor{solarized-base00} {RGB}{101, 123, 131}
\definecolor{solarized-base0}  {RGB}{131, 148, 150}
\definecolor{solarized-base1}  {RGB}{147, 161, 161}
\definecolor{solarized-base2}  {RGB}{238, 232, 213}
\definecolor{solarized-base3}  {RGB}{253, 246, 227}
\definecolor{solarized-yellow} {RGB}{181, 137, 000}
\definecolor{solarized-orange} {RGB}{203, 075, 022}
\definecolor{solarized-red}    {RGB}{220, 050, 047}
\definecolor{solarized-magenta}{RGB}{211, 054, 130}
\definecolor{solarized-violet} {RGB}{108, 113, 196}
\definecolor{solarized-blue}   {RGB}{038, 139, 210}
\definecolor{solarized-cyan}   {RGB}{042, 161, 152}
\definecolor{solarized-green}  {RGB}{133, 153, 000}

\newtheorem{definition}{Definition}

%%%%%% for algorithms
\errorcontextlines\maxdimen
\algrenewcommand\alglinenumber[1]{\tiny\color{solarized-base0} #1}

% begin vertical rule patch for algorithmicx (http://tex.stackexchange.com/questions/144840/vertical-loop-block-lines-in-algorithmicx-with-noend-option)
\makeatletter
% start with some helper code
% This is the vertical rule that is inserted
    \newcommand*{\algrule}[1][\algorithmicindent]{\makebox[#1][l]{\color{solarized-base0}\hspace*{.5em}\thealgruleextra\vrule height \thealgruleheight depth \thealgruledepth}}%
% its height and depth need to be adjustable
\newcommand*{\thealgruleextra}{}
\newcommand*{\thealgruleheight}{.75\baselineskip}
\newcommand*{\thealgruledepth}{.25\baselineskip}

\newcount\ALG@printindent@tempcnta
\def\ALG@printindent{%
    \ifnum \theALG@nested>0% is there anything to print
        \ifx\ALG@text\ALG@x@notext% is this an end group without any text?
            % do nothing
        \else
            \unskip
            \addvspace{-1pt}% FUDGE to make the rules line up
            % draw a rule for each indent level
            \ALG@printindent@tempcnta=1
            \loop
                \algrule[\csname ALG@ind@\the\ALG@printindent@tempcnta\endcsname]%
                \advance \ALG@printindent@tempcnta 1
            \ifnum \ALG@printindent@tempcnta<\numexpr\theALG@nested+1\relax% can't do <=, so add one to RHS and use < instead
            \repeat
        \fi
    \fi
    }%
\usepackage{etoolbox}
% the following line injects our new indent handling code in place of the default spacing
\patchcmd{\ALG@doentity}{\noindent\hskip\ALG@tlm}{\ALG@printindent}{}{\errmessage{failed to patch}}
\makeatother

% the required height and depth are set by measuring the content to be shown
% this means that the content is processed twice
\newbox\statebox
\newcommand{\myState}[1]{%
    \setbox\statebox=\vbox{#1}%
    \edef\thealgruleheight{\dimexpr \the\ht\statebox+1pt\relax}%
    \edef\thealgruledepth{\dimexpr \the\dp\statebox+1pt\relax}%
    \ifdim\thealgruleheight<.75\baselineskip
        \def\thealgruleheight{\dimexpr .75\baselineskip+1pt\relax}%
    \fi
    \ifdim\thealgruledepth<.25\baselineskip
        \def\thealgruledepth{\dimexpr .25\baselineskip+1pt\relax}%
    \fi
    %\showboxdepth=100
    %\showboxbreadth=100
    %\showbox\statebox
    \State #1%
    %\State \usebox\statebox
    %\State \unvbox\statebox
    %reset in case the next command is not wrapped in \myState
    \def\thealgruleheight{\dimexpr .75\baselineskip+1pt\relax}%
    \def\thealgruledepth{\dimexpr .25\baselineskip+1pt\relax}%
}
% end vertical rule patch for algorithmicx

%%%% For listings (code blocks)
\lstset{
    basicstyle={\footnotesize\ttfamily},
    numbers=left,
    keywordstyle=\color{solarized-orange}\bfseries,
%    identifierstyle=\color{solarized-base02},
    stringstyle=\color{solarized-cyan},
    commentstyle=\color{solarized-base1}\itshape,
    numberstyle=\tiny\color{solarized-base0},
    columns=flexible,
    stepnumber=1,
    numbersep=5pt,
    backgroundcolor=\color{solarized-base3},
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2,
    captionpos=b,
    breaklines=true,
    breakatwhitespace=true,
    breakautoindent=true,
    escapeinside={\%*}{*)},
    linewidth=\textwidth,
    basewidth=0.5em,
}

%\setlength{\belowcaptionskip}{-2pt}
\setlength{\abovecaptionskip}{-12pt}

%%%% For links URL
\hypersetup{colorlinks=true, linkcolor=solarized-blue, citecolor=solarized-base1, filecolor=solarized-magenta, urlcolor=solarized-cyan}

% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\begin{document}
\maketitle
\begin{abstract}
As of recent years, automated planning domain has mainly focused on
performances and advances in state space planning to improve
scalability. That orientation shadows other less efficient ways of doing
like Partial Ordering Planning (POP) that has the advantage to be much
more flexible. This approach generates plans that can be refined easily
and can be used to achieve greater resilience, repairing capabilities
and soft resolution that applications based on online plan recognition
or decision-making can benefit from. This paper presents a set of
algorithms, named solving and Defect Aware Partial Ordering Planning
(SODA POP), that aims at targeting these objectives and maintain a high
plan quality. Our algorithms create proper offline plans for goals and
use an effective defect fixing algorithm to repair input plans even if
these plans are corrupted. SODA POP can also use healer actions and
links in order to always return a valid plan even in cases where the
problem is impossible to solve using problem derivation. Some relevant
properties of these algorithms are analyzed in this article and
experimental results show interesting performances for online planning
and repairing.
\end{abstract}

\section*{Introduction}\label{introduction}
\addcontentsline{toc}{section}{Introduction}

For some time, Partial Order Planning (POP) has been the most popular
approach to planning resolution. This kind of algorithms are based on
\emph{least commitment strategy} on plan step ordering that allow
actions to be flexibly interleaved during execution {[}1{]}. Thus, the
way the search is made using flexible partial plan as a search space
allowed for more versatility for a wide variety of uses. As of more
recent years, new state space search approach and heuristics {[}2{]}
have been demonstrated to be more efficient than POP planners due to the
simplicity and relevance of state representations opposed to partial
plans {[}3{]}. This have made the search of performance the main axis of
research for planning.

While this goal is justified, it shadows other problems that some
practical applications cause. For example, the flexibility of Plan Space
Planning (PSP) is an obvious advantage for applications needing online
planning: plans can be repaired on the fly as new information and goals
are considered. The idea of using POP for online planning and repairing
plans instead of replanning from scratch is not new {[}4{]}, but has
never been paired with the resilience that some other cognitive
applications may need, especially when dealing with sensors noise in
input data. This resilience makes fixing errors easier than with an
external algorithm as the plan logic allows for context driven decision
on the way to repair the issues. For example, if an action becomes
unrelevant or incoherent, the flaw in the partial plan makes the issue
explicit and therefore easier to fix. Applications might sometimes
provide plans that can contain errors and imperfections that will
decrease significantly the efficiency of the computation and the quality
of the output plan. Additionnaly, these plans may become totally
unsolvable. This problem is to our knowledge not treated in planning of
all forms (state planning, PSP, and even constraint satisfaction
planning) as usually the aim is to find a solution relative to the
original plan (which makes sense). However, as we proceed a mechanism of
\emph{problem derivation} may be required. This will allow soft solving
of any problem regardless of its resolvability.

One of the applications that needs these improvements is plan
recognition with the particular use of off-the-shelf planners to infer
the pursued goal of an agent where online planning and resilience is
particularly important {[}5{]}. This method adds dummy actions that need
to be satisfied by modifying the goal, to ensure that the observed plan
is selected by the planner. A system that is able to repair plan will
ease such an application. Another application is decision-making in
dynamic environments. Indeed, having a plan that details all steps and
explicits the ones that are not possible, can help decide on the plan of
actions to take. These problems call for new ways to improve the answer
of a planner. These improvements must provide relevant plan information
pointing out exactly what needs to be done in order to make a planning
problem solvable, even in the case of obviously inconsistent input data.
This paper aims to solve this while preserving the advantages of PSP
algorithms (flexibility, easy fixing of plans, soundness and
completeness). We propose a new set of auxiliary algorithms that
combined forms our SOft solving and Defect Aware Partial Ordering
Planning (SODA POP) algorithm that targets those issues and allows
making POP algorithm more resilient, efficient and relevant. This is
done by pre-emptively computing proper plans for goals, by solving new
kinds of defects that input plans may exhibit, and by healing
compromised plan by deriving the initial problem with forged actions to
allow resolution. To explain this work we first introduce a few notions,
notations and specificities of existing POP algorithms. Then we present
and illustrate their limitations, categorizing the different defects
arising from the resilience problem and explaining our auxiliary
algorithms, their uses and properties. We compare the performance,
resilience and quality of POP and our solution.

\section{Definitions}\label{definitions}

First, we introduce some definitions and notations for mathematical
representation to present our model.

\subsection{Classical planning}\label{classical-planning}

\begin{definition}[Fluents]

A fluent is a property of the world. We note \(\lnot f\) the
complementary fluent of \(f\) meaning that \(f\) is true when
\(\lnot f\) is false and vice-versa.

\end{definition}

It is often represented by first order logical propositions but in this
paper we choose to focus on the algorithm and to represent fluents as
simple literals (fully instantiated) using \(\mathbb{Q}^*\), the set of
relative integers without \(0\), as the fluent domain. We use negative
integers to represent opposite fluents.

\begin{definition}[State]

A state is defined as a set of fluents. States can be additively
combined. We note
\(s_1 + s_2 = \left( s_1 \cup s_2 \right) - \left\{ f \middle| f \in s_1 \land \lnot f \in s_2 \right\}\)
such operation. It is the union of the fluents with an erasure of the
complementary ones.

\end{definition}

\begin{definition}[Action]

An action is a state operator. It is represented as a tuple
\(a = \langle pre, eff \rangle\) with \(pre\) and \(eff\) being sets of
fluents, respectively the preconditions and the effects of \(a\). An
action can be used only in a state that verifies its preconditions. We
note \(s \models a \Leftrightarrow pre(a) \subseteq s\) the verification
of an action \(a\) by a state \(s\).

\end{definition}

An action \(a\) can be functionally applied to a state \(s\) as follows:
\[a:= \substack{ \left\{ s \models a \middle| s \in S\right\} \to S\\
    a(s) \mapsto s + eff(a)}\] with \(S\) being the set of all states.
This means that an action adds all its effects to the state of the world
on application (by removing complementary fluents if needed). We
distinguish between two specific kinds of actions: actions with no
preconditions are synonymous to a state and those with empty effect are
called a goal.

\subsection{Plan Space Planning}\label{plan-space-planning}

\subsubsection{Problem}\label{problem}

We define a partial plan satisfaction problem as a tuple noted
\(P = \langle A, I, G, p \rangle\) with:

\begin{itemize}
\tightlist
\item
  \(I\) and \(G\) being the pseudo actions representing respectively the
  initial state and the goal.
\item
  \(p\) being a partial plan to refine.
\item
  \(A\) the set of all actions.
\end{itemize}

\begin{definition}[Partial Plan]

A \emph{partial plan} is a tuple \(p = \langle A_p, L\rangle\) where
\(A_p\) is a set of steps (actions) and \(L\) is a set of causal links
of the form \(a_i \xrightarrow{f} a_j\), such that
\(\{ a_i, a_j \} \subseteq A_p \land f \in eff(a_i) \cap pre(a_j)\) or
said otherwise, this causal link means that the fluent \(f\) is provided
by an effect of \(a_i\) to a precondition of \(a_j\). We include the
ordering constraints of PSP in the causal links. An ordering constraint
is noted \(a_i \to a_j\) and means that the plan means \(a_i\) as a step
that is prior to \(a_j\) without specific cause (usually because of
threat resolution). We note the order relation over \(A\) for an action
\(a_i\) that is prior to the action \(a_j\) following all the causal
links and ordering constraints \(a_i \succ a_j\). We can consider it as
a graph formed by step as vertices and causal links as edges.

\end{definition}

The figure \ref{fig:legend} details how the elements of partial plans
are represented in the rest of this paper.

\begin{figure}[htbp]
\centering
\includegraphics{graphics/legend.pdf}
\caption{Global legend for partial plans\label{fig:legend}}
\end{figure}

In this figure the fluent \(6\) is a precondition of the action \(a\)
and also an effect. The causal link is providing the fluent \(6\). The
dotted line is an ordering constraint specifying which action should be
placed before another. \(I\) and \(G\) are respectively the initial and
goal step.

\subsubsection{Flaws}\label{flaws}

When refining a partial plan, we need to fix flaws. Those could be
present in the input or created by the refining process. Flaws can
either be unsatisfied subgoals or threats to causal links.

\begin{definition}[Subgoal]

A subgoal \(s\) is a precondition of an action \(a_s \in A_p\) with
\(s \in pre(a_s)\) that isn't satisfied by any causal link. We can note
a subgoal as:
\[a_i \xrightarrow{s} a_s \notin L \mid \{ a_i, a_s \} \subseteq A_p \]

\end{definition}

\begin{definition}[Threat]

A step \(a_t\) is said to threaten a causal link
\(a_i \xrightarrow{t} a_j\) if and only if
\[\neg t \in eff(a_t) \land a_i \succ a_t \succ a_j \models L\]

\end{definition}

In other words, the action has a complementary effect that can be
inserted between two actions needing this fluent while being consistent
with the ordering constraint in \(L\).

\begin{definition}[Resolvers]

Resolvers are a set of actions and causal links
\(r = \langle A_r , L_r \rangle\) that fixes flaws.

\begin{itemize}
\tightlist
\item
  A resolver for a subgoal is an action \(a_r \in A\) that has \(s\) as
  an effect \(s \in eff(a_r)\) inserted along with a causal link noted
  \(a_r \xrightarrow{s} a_s\).
\item
  The usual resolvers for a threat are either \(a_t \to a_i\) or
  \(a_j \to a_t\) which are called respectively promotion and demotion
  links.
\end{itemize}

\end{definition}

\begin{definition}[Solution]

The solution of a PSP problem is a valid partial plan that respects the
specification of said problem (only using actions in \(A\) and having
the correct initial and goal step).

\end{definition}

\begin{definition}[Consistency]

A partial plan is consistent if it contains no ordering cycles. The
directed graph formed by step as vertices and causal links as edges
isn't cyclical. This is important to guarantee the soundness.

\end{definition}

\begin{definition}[Flat Plan]

We can instantiate one or several flat plans from a partial plan. A flat
plan is an ordered sequence of actions \(\pi = [ a_1, a_2 \ldots a_n]\)
that acts like a action \(\pi = \langle pre_\pi, eff_\pi \rangle\) and
can be applied to a state \(s\) using functional composition operation
\(\pi := \bigcirc_{i=1}^n a_n\).

\end{definition}

We call a flat plan valid if and only if it can be functionally applied
on an empty state. We note that this is different from classic state
planning because in our case the initial state is the first action that
is already included in the plan.

\begin{definition}[Validity]

A partial plan is valid if and only if it is consistent and if all flat
plans that can be generated are valid.

\end{definition}

\section{Classical POP}\label{classical-pop}

Partial Order Planning (POP) is a popular implementation of the general
PSP algorithm. It refines a partial plan by trying to fix its flaws and
is proven to be sound and complete {[}6{]}.

\begin{algorithm}\caption{Classical Partial Order Planning}\label{pop}\begin{algorithmic}[1]

\footnotesize
\Function{pop}{Queue of Flaws $agenda$, Problem $P$}
\State \Call{populate}{$agenda$, $P$} \Comment{Only on first call}
\If{$agenda = \emptyset$} \State \Return Success
\Comment{Stop all recursion} \EndIf
    \State Flaw \(f\gets agenda.popFromQueue\)
\Comment{Stop all recursion} \Comment{First element of the queue}
\State Resolvers \(R \gets\) \Call{resolvers}{$f$, $P$}
\State \Comment{Ordered resolvers to try} \ForAll{$r \in R$}
\Comment{Non deterministic choice operator}
\State \Call{apply}{$r$, $P.p$} \Comment{Apply resolver to partial plan}
\If{\Call{consistent}{$P.p$}} \Comment{$P.p$ is the partial plan}
\State \Call{pop}{$agenda \cup$ \protect\Call{relatedFlaws}{$f$}, $P$}
\State \Comment{Finding new flaws introduced by the resolver} \Else 
            \State \Call{revert}{$r$, $P.p$}
\Comment{Undo resolver insertion} \EndIf
    \EndFor
    \State \Return Failure
\State \Comment{Revert to last non deterministic choice of resolver}
\EndFunction

\end{algorithmic}\end{algorithm}

Algorithm 1 presents the base algorithm for a planer in the plan space.
POP implementation uses an agenda of flaws that is efficiently updated
after each refinement of the plan. A flaw is selected for resolution and
we use a non deterministic choice operator to pick a resolver for the
flaw. The resolver is inserted in the plan and we recursively call the
algorithm on the new plan. On failure, we return to the last non
deterministic choice to pick another resolver. The algorithm ends when
the agenda is empty or when there is no more resolver to pick for a
given flaw. Before continuing, we present a simple example of classical
POP execution with the following example of a problem.

\begin{itemize}
\tightlist
\item
  An initial state \(I = \langle \emptyset , \{ 1, 2 \} \rangle\) and a
  goal \(G = \langle \{ 3, 4, -5, 6 \}, \emptyset \rangle\) encoded as
  dummy steps.
\item
  \(a \langle \{1\}, \{3,5\} \rangle\),
  \(b \langle \{2\}, \{4\} \rangle\) and
  \(c \langle \{5\}, \{6\} \rangle\) are simple actions that are useful
  to achieve the goal.
\item
  \(n \langle \{-8, 7\}, \{-7, 8\} \rangle\) and
  \(l \langle \{-7, 8\}, \{-8, 7\} \rangle\) are looping actions that
  are meant to cause cycles.
\item
  \(t \langle \{4\}, \{-5\} \rangle\) is meant to be threatening to the
  plan's integrity and will generate threats.
\item
  \(u \langle \{4\}, \emptyset \rangle\),
  \(v \langle \{4\}, \{4\} \rangle\),
  \(w \langle \{9\}, \{3\} \rangle\),
  \(x \langle \{7\}, \{-2,2\} \rangle\) are actions that will cause
  problems. \protect\hyperlink{defects}{(cf.~Defects section)}.
\end{itemize}

This standard way of doing has seen multiple improvements over
expressiveness like with UCPOP {[}7{]}, Hierarchical Tasks Network to
add more user control over sub-plans {[}8{]}, cognition with defeasible
reasoning {[}9{]}, or speed with multiple ways to implement the popular
fast forward method from state planning {[}10{]}. However, all these
variants do not treat the problem of online planning, resilience and
soft solving. Some other closer works like {[}4{]} treat the problem of
online planning by identifying and removing trees of the partial plan.
This is done with a heuristic that choses an unrefinemment strategies
that causes a heavy replanning of the problem even if only one action
needs removal. This is a significant problem when trying, for example,
to adapt a plan with minimal changes due to replanning. Indeed, all
these problems can affect POP's performance and quality.

\begin{figure}[htbp]
\centering
\includegraphics{graphics/pop.pdf}
\caption{Standard POP result to the problem\label{fig:pop}}
\end{figure}

This example has been crafted to illustrate problems with standard POP
implementations. We give a possible resulting plan of standard POP in
figure \ref{fig:pop}. There are some issues as for how the plan has been
built. The action \(v\) is being used even if it is useless since \(b\)
already provided fluent \(4\). We can also notice that despite being
contradictory the action \(x\) raised no alarm. As for ordering
constraints, we can clearly see that the link \(a \to t\) is redundant
with the path \(a \xrightarrow{5} c \to t\) that already ensures that
constraint by transitivity. Also, some problems arise during execution
with the selection of \(w\) that causes a dead-end.

Of course the flaw selection mechanism of certain variants can prevent
that to happen in that case. But often flaw selection mechanisms are
more speed oriented and will do little if a toxic action seems to fit
better than a more coherent but complex one. All these issues are caused
by what we call \emph{defects} as they are not regular PSP flaws but
they still cause problems with execution and results. We will address
these defects and propose a way to fix them in
\protect\hyperlink{defects}{the next section}.

\section{Auxiliary algorithms to POP}\label{auxiliary-algorithms-to-pop}

In order to improve POP algorithms' resilience, online performance and
plan quality, we propose a set of auxiliary algorithms that provides POP
with a clean and efficiently populated initial plan. The complete
algorithm will be presented in \protect\hyperlink{soda}{the next
section} as a combination of all auxiliary algorithms and classical POP.

\subsection{Proper plan generation}\label{proper-plan-generation}

\begin{algorithm}\caption{Proper plan generation for a given goal $g$}\label{properplan}\begin{algorithmic}[1]

\footnotesize
\Function{properPlan}{Goal $g$, Actions $A$} \State Partial Plan
\(p \gets \emptyset\) \State Actions \(relevants \gets\)
\Call{satisfy}{$g$, $A$, $p$}
\State \Comment{Satisfy the goal with all participating actions and links}
\State Queue of Actions \(open \gets relevants\)
\While{$open \neq \emptyset$} \State Action \(a\gets open.popFromQueue\)
\State Actions \(candidates \gets\) \Call{satisfy}{$a$, $A$, $p$}
\ForAll{$candidate \in candidates$} \If{$candidate \notin relevants$}
\State \(open.pushToQueue(candidate)\) \EndIf
        \EndFor
    \EndWhile
\EndFunction

\end{algorithmic}\end{algorithm}

\begin{figure}[htbp]
\centering
\includegraphics{graphics/proper.pdf}
\caption{Proper plan of the example goal\label{fig:proper}}
\end{figure}

As in online planning goals can be known in advance, we propose a new
mechanism that generates proper plans for goals. We take advantage of
the fact that this step can be done offline to improve the performance
of online planning. This offline execution prevents us to access the
details of the initial state of the world as it will be defined at
runtime. We define for that the concept of \emph{participating action}.
An action \(a \in A\) participates in a goal \(G\) if and only if \(a\)
has an effect \(f\) that is needed to accomplish \(G\) or that is needed
to accomplish another participating action's preconditions. A proper
plan is a partial plan that contains all participating actions as steps
and causal links that bind them with the step they are participating in.
This proper plan is independent from the initial step because we might
not have the initial step at the time of the proper plan generation.
Once applied on the previous example, proper plan generation returns the
partial plan presented in figure \ref{fig:proper}. This partial plan
doesn't have initial state because of its offline nature. It also shows
several cycles and obvious problems. However, it has all the steps of
the correct final plan. This algorithm helps the POP algorithm by
prefetching all the actions subgoals might need. However, this result
needs to be cleaned first to be helpful.

\hypertarget{defects}{\subsection{Defect resolution}\label{defects}}

\begin{algorithm}\caption{Defect resolution}\label{defectresolution}\begin{algorithmic}[1]

\footnotesize
\Function{clean}{Problem $P$}
\State \Call{assert}{$pre(P.I) = \emptyset$}
\State \Call{assert}{$eff(P.G) = \emptyset$}
\State \(P.p.A_p \gets P.p.A_p \cup \{P.I, P.G\}\)
\State \(defects \gets\) \Call{findDefects}{$P$}
\State \Call{fix}{$defects$, $P$} \EndFunction

\Function{fix}{Defects $defects$, Problem $P$}
\ForAll{Defect $d \in defects$} \State \(defect.\)\Call{fix}{$P$}
\Comment{Fixing is proper to the defect} \EndFor
\EndFunction

\Function{findDefects}{Problem $P$} \State \Return \Call{illegal}{$P$}
\(+\) \Call{interfering}{$P$}
\State \Comment{Concatenate the found illegal and interfering defects}
\EndFunction

\end{algorithmic}\end{algorithm}

When the POP algorithm is used to refine a given plan (that was not
generated with POP or that was altered), a set of new defects can be
present in it interfering in the resolution and sometimes making it
impossible to solve. We emphasize that these \emph{defects are not
regular POP flaws} but new problems that classical POP can't solve. The
aim of this auxiliary algorithm is to clean the plans from such defects
in order to improve resilience and plan quality.

\begin{definition}[Plan Quality]

Plan quality is an indicator that is measured by the number of defects
in a partial plan. There are two specific indicators for a plan quality:

\begin{itemize}
\tightlist
\item
  Action quality is the number of defects related to actions divided by
  the number of actions.
  \[quality_{action}(p) =  1 - \left( |ActionDefects(p)| \over |p.A_p| \right)\]
\item
  Link quality is the number of defects related to links divided by the
  number of links.
\end{itemize}

\end{definition}

From there it is obvious that plan quality will improve over POP since
SODA POP algorithm is guaranteed to remove all defects in a plan. There
are two kinds of defects: the illegal defects that violate base
hypothesis of POP and the interference defects that can lead to
excessive computational time and to poor plan quality.

\subsubsection{Illegal defects}\label{illegal-defects}

These defects are usually hypothesized out by existing algorithms. They
are illegal use of partial plan representation and shouldn't happen
under regular circumstances. They may appear if the input is generated
by an approximate cognitive model that doesn't ensure consistency of the
output or by unchecked corrupted data. These defects will most of the
time simply break regular POP algorithms or at least make the
performances decrease significantly. Illegal defects and solutions to
fix them are presented in the following section. The first of these
defects is pre-existing cycles in the input plan.

\begin{definition}[Cycles]

A cycle is the corresponding notion as in graph theory. Let's consider
the graph of a partial plan. A cycle in this graph corresponds to a
cycle of actions in the partial plan.

\end{definition}

A plan cannot contain cycles as it makes it impossible to complete.
Cycles are usually detected as they are inserted in a plan but poor
input can potentially contain them and break the POP algorithm as it
cannot undo cycles. We use strongly connected component detection
algorithm to detect cycles. Upon cycle detection, the algorithm can
remove arbitrarily a link in the cycle to break it. In our example the
actions \(n\) and \(l\) are meant to cause such cycles. In a plan some
actions can be illegal for POP. Those are the actions that are
inconsistent.

\begin{definition}[Inconsistent actions]

An action \(a\) is contradictory if and only if
\[\exists f \{f, \lnot f \} \in eff(a) \lor \{f, \lnot f \} \in pre(a)\]

\end{definition}

We remove only one of those effects or preconditions based on the usage
of the action in the rest of the plan. If none of those are used, we
choose to remove both. In our previous example, the action \(x\) is one
of these inconsistent actions with fluent \(2\) and \(-2\) in its
effects.

\begin{definition}[Toxic actions]

Toxic actions are those that have effects that are already in their
preconditions or empty effects. They are defined as:
\[a | pre(a) \cap eff(a) \neq \emptyset \lor eff(a) = \emptyset\]

\end{definition}

Toxic actions can damage a plan as well as make the execution of POP
algorithm much longer than necessary. This is fixed by removing the
toxic fluents (\(pre(a) \nsubseteq eff(a)\)) by updating the effects
with \(eff(a) = eff(a)-pre(a)\). If the effects become empty, the action
is removed altogether from plan steps and \(A\). In our previous example
the actions \(u\) and \(v\) are toxic as the fluent \(4\) is in the
effects and preconditions of \(v\), and \(u\) has empty effects. In this
case, these actions will be removed by the algorithm as they don't have
any other effects.

The defects can be related to incorrect links. The first of which are
liar links.

\begin{definition}[Liar links]

A liar link is a link that doesn't hold a fluent in the preconditions or
effect of its source and target. We can note:
\[a_i \xrightarrow{f} a_j | f \notin eff(a_i) \cap pre(a_j)\]

\end{definition}

A liar link can be either already present in the data or created by the
removal of an effect of an inconsistent or toxic action (with the causal
link still remaining).

We call lies, fluents that is held by links without being in the
connected actions . To resolve the problem we remove all lies and add
all savior fluents, i.e.~a fluent in \(eff(a_i) \cap pre(a_j)\) that
isn't already held by the link. We delete the link all together if it
doesn't link any fluent.

\subsubsection{Interference defects}\label{interference-defects}

This kind of defects is not as problematic as the illegal ones: they
won't make the plan unsolvable but they can still cause performance and
quality drops. These defects can happen in POP generated plans to some
extends. Interference defects and solutions to fix them are presented in
the following.

\begin{definition}[Redundant links]

A redundant link is an ordering constraint (causal link without fluents)
with another equivalent path of longer length int the graph form of the
partial plan. It means that an ordering constraint \(a_i \to a_j\) is
redundant if and only if it exists another path from \(a_i\) to \(a_j\)
of length greater to \(1\).

\end{definition}

Since POP relies on those additional links, this part focuses on
removing the ones that were generated for threat removal purpose to
simplify the plan. For example, in figure \ref{fig:pop} we can see that
the ordering constraint from \(a\) to \(t\) is redundant with the path
\(a \xrightarrow{5} c \to t\) in that, it isn't needed to resolve the
threat. Therefore, the algorithm would remove it. This reduces the
number of edges in the plan and therefore simplifies it.

Causal links can be found to compete with one another.

\begin{definition}[Competing causal links]

A competing link \(a_i \xrightarrow{f} a_k\) competes with another link
\(a_j \xrightarrow{f} a_k\) if it provides the same fluent to the same
action.

\end{definition}

The main difference with redundant links is that competing links are not
related to threat resolution generated ordering constraint but confronts
two causal links that can provide the same fluent. Another difference is
that it cannot happen in classical POP algorithm. This kind of defects
is making the plan more complex and can also hide useless actions which
would potentially cause more flaws to handle than necessary. In order to
remove the least interesting link in competing causal links, we need to
compare their respective providing actions, to chose the most useful
ones.

\begin{definition}[Usefulness of an action]

The usefulness of an action is the number of participating links
(outgoing causal links) divided by the number of needing links (incoming
causal links). If there are no needing links, the usefulness is simply
the number of participating links.

In the special cases if the action is either an initial step or a goal,
the usefulness is \(+\infty\). The usefulness of savior action is \(0\).

\end{definition}

We choose the link with the source action having the highest usefulness.
This replaces unnecessary savior actions and allows reducing the
usefulness of the least action in order to eventually make it an orphan
and prune it from the plan.

\begin{figure}[htbp]
\centering
\includegraphics{graphics/competing.pdf}
\caption{Example of choice for cutting competing links based on
usefulness of actions\label{fig:competing}}
\end{figure}

On the left example of figure \ref{fig:competing}, the action \(a_j\)
participates much more than the action \(a_i\) and therefore the link to
be removed would be \(a_i \xrightarrow{f} a_k\). On the right example,
the actions don't have different outgoing links but the action \(a_j\)
is here much needier than its competitor. Therefore, the link to be
removed is the link \(a_j \xrightarrow{f} a_k\).

Actions can sometimes have no use in a plan as they don't contribute to
it.

\begin{definition}[Orphan actions]

Orphan actions are actions without any links or actions with no outgoing
path to the goal (meaning that it doesn't participate in the plan). That
also concerns actions that would become orphans if another orphan action
is removed.

\end{definition}

\begin{algorithm}\caption{Orphan actions finding}\label{orphanaction_find}\begin{algorithmic}[1]

\footnotesize
\Function{$OrphanAction.$find}{Problem $P$} \State Actions
\(orphan \gets \emptyset\) \State int \(oldSize\) \Repeat
        \State \(oldSize \gets |orphan|\)
\ForAll{Action $step \in P.p.A_p$} \If{$step \in orphan$}
\State continue \EndIf
            \State int \(towardOrphan \gets 0\) \State Causal Links
\(outgoing \gets\) \State \Call{outgoingEdgesOf}{$step$, $P.p$}
\ForAll{Causal Link $link \in outgoing$} \If{$link.target \in orphan$}
\State \(towardOrphan ++\) \EndIf
            \EndFor
            \If{$|outgoing| = towardOrphan \land step \notin \{ P.I, P.G\}$}
\State \(orphan \gets orphan \cup \{step\}\) \EndIf
        \EndFor
    \Until{$oldSize \neq |orphan|$} \State \Return \(orphan\)
\EndFunction

\end{algorithmic}\end{algorithm}

In order to fix this, we derive the idea of the backward removal tree of
{[}4{]}. Algorithm 4, removes an action if all its outgoing causal links
are leading to useless actions. It iterates over all actions until no
new useless action is discovered. All useless actions are then removed
from the plan.

\begin{definition}[Competing actions]

In the same way, links can be competing toward a common needer,
sometimes some actions that are more suited to achieve a goal than
others. These actions are taken into account if they have effects that
are carried by the link, if they are more useful than the source of the
existing link and if they wouldn't cause a cycle.

\end{definition}

In such a case the link is removed and another one is formed by the
better suited action.

\subsection{Soft resolution}\label{soft-resolution}

\begin{algorithm}\caption{Soft resolution healing}\label{softresolution}\begin{algorithmic}[1]

\footnotesize
\Function{heal}{Problem $P$} \State int \(minViolation \gets \infty\)
\State Plan \(best \gets P.p\) \State Flaw \(annoyer\)
\ForAll{$\langle flaw, plan \rangle \in P.partialSolutions$} \State int
\(currentViolation \gets\) \Call{violation}{$plan$, $P.G$}
\If{$currentViolation < minViolation$} \State \(best \gets plan\)
\State \(annoyer \gets flaw\)
\State \(minViolation \gets currentViolation\) \EndIf
    \EndFor
    \State \(P.p \gets best\)
\State \(P.partialSolutions \gets \emptyset\)
\ForAll{Resolver $resolver \in$ \Call{healers}{$annoyer$}}
\State \Call{apply}{$resolver$, $P.p$} \EndFor
\EndFunction

\end{algorithmic}\end{algorithm}

This auxiliary algorithm is meant to deal with failure. It heals the
plan to make the failure recoverable for the next iteration of POP. Of
course it can't fix the plan by keeping the problem as it is. This
obviously breaks some properties as the algorithm no longer adheres to
the specification of the input, but in exchange it will always issue a
valid plan whatever happens.
\protect\hyperlink{hypersoundness}{(cf.~Hypersoundness section)}.

Soft failure is useful when the precision and validity of the output is
not the most important criteria we look for. In some cases (like in
recognition processes) it is more useful to have an output even if it is
not exact than no output at all. That is why we propose a soft failing
mechanism for the POP algorithm. Following, we define first some
notions, then we explain the healing algorithm.

\begin{definition}[Needer]

A needer is an action that needs a resolution related to a flaw. We
define different types of needers according to the type of the flaw.

\begin{itemize}
\item
  For a subgoal \(a_s \xrightarrow{s} a_n\) the needer is the action
  \(a_n\) that has an unsatisfied precondition provided by an eventual
  action \(a_s\) in the current partial plan.
\item
  For a threat \(a_t\) of a link \(a_p \xrightarrow{t} a_n\) the needer
  is the target \(a_n\) of the threatened causal link.
\end{itemize}

\end{definition}

\begin{definition}[Proper fluents]

A proper fluent of a flaw is the one that caused the flaw. For a subgoal
\(a_n \xrightarrow{s} a_s\), it is the unsatisfied precondition \(s\).
For a threat \(a_t\) of a causal link \(a_p \xrightarrow{t} a_n\), it is
the fluent \(t\) held by the threatened causal link.

\end{definition}

\begin{definition}[Savior]

The savior of a flaw is the forged action
\(a_s = \langle \emptyset, \{p\} \rangle | a_s \notin A\) with \(p\)
being the proper fluent of the flaw.

\end{definition}

The concept of healer is made to target rogue flaws that caused total
failure.

\begin{definition}[Healers]

A healer is a resolver built using the savior of the flaw in order to
provide the missing fluents to it. The general formula of a healer is
\(a_s \xrightarrow{p} a_n\) with \(a_s\) being the savior of the flaw.

\end{definition}

For threats, we need an additional healer specified as an ordering
constraint from the threatening action to the savior \(a_t \to a_s\) to
ensure that the savior acts after the threat and therefore provides the
proper fluent for the needer.

\begin{definition}[Violation degree]

The violation degree \(v(p)\) of a plan \(p\) is an indicator of the
violation regarding the initial problem. We note
\(v(p) = |flaws(p)| + |saviors(A_p)|\). Said otherwise, it is the sum of
the number of flaws and the number of saviors in the plan.

\end{definition}

\subsubsection{Healing process}\label{healing-process}

The healing method is to keep track of reversions in the algorithm by
storing the partial plan and the unsatisfiable flaw each time a non
deterministic choice fails. We note the set of these failed plans \(F\).
As the POP algorithm encounters a final failure, this auxiliary
algorithm get called. The aim is to evaluate each backtracked partial
plan to choose the best one.

Therefore, we add an order relation for \(F\) noted
\[\prec : p \prec q \iff v(p) < v(q) | \{p, q\} \subseteq F\]

Once the POP algorithm fails completely, the soft failing algorithm can
be invoked to heal the plan. It chooses the best plan
\(b | \forall p \in F, b \prec p\) to heal it. If two plans have the
same violation degree, the algorithm chooses the first one to have
happened. The healing process is similar to how POP works: we apply the
healer of the flaw that caused the failure of the partial plan we
chosen. We empty the set \(F\) to allow POP to iterate further since the
flaw is resolved. The healing process can be done for each unsolved
flaws as POP fails repeatedly. This ensures some interesting properties
explained in the following section.

\hypertarget{soda}{\section{SODA POP and its properties}\label{soda}}

The combination of classical POP and all the auxiliary algorithms
presented previously gives SODA POP algorithm detailed in Algorithm 6.
In this section, we focus on the properties of SODA POP.

\begin{algorithm}\caption{SODA POP}\label{soda}\begin{algorithmic}[1]

\footnotesize
\Function{soda}{Problem $P$} \State \(P.p \gets\)
\Call{properPlan}{$P.G$, $P.A$} \Comment Offline execution
\State \Call{clean}{$P$} \State bool \(valid \gets false\)
\While{$\lnot valid$} \State \(valid \gets\) \Call{pop}{$P$}
\(= Success\) \If{$valid$} \State \Call{clean}{$P$}
\State \Return \(Success\) \EndIf
        \State \Call{heal}{$P$} \EndWhile
\EndFunction

\end{algorithmic}\end{algorithm}

\subsection{Convergence property of POP}\label{convergence}

First, we prove the convergence of POP. The classic planning problem is
already proven to be decidable without functions in the fluents {[}3{]}.
Therefore, we can categorize the termination cases.

\begin{proof}[Proof of convergence]

In the case of a solvable problem, POP is proven to be complete. This
ensures that it will find a solution for the problem and therefore
terminate. Let \(flaws(p)\) be the set of flaws of a given partial plan.
The number of flaws is the number of subgoals plus the number of threats
since \(flaws(p) = subgoals(p) \cup threats(p)\). We consider the number
of actions \(|A|\) as being finite. Therefore, the number of steps in
the plan is at worse \(|A_p| = |A|\). We also assume that actions have a
finite number of preconditions and effects (since we don't use functions
over fluents). This leads to
\(|subgoals(p)| < \sum_{a \in A_p} |pre(a)| < \infty\) and
\(|threats(p)| < |L| \le \sum_{a \in A_p} |pre(a)| < \infty\).

This means the number of all possible flaws is finite. As POP resolves
these flaws it will decrease their number and iterate over resolvers.
The number of resolvers is \(|subgoals(p)| * |A| + |threats(p) * 2\) and
is also finite. This means that the iteration will in the worst case be
equal to the number of resolvers before failing. This proves termination
and therefore that POP converges. \qedhere

\end{proof}

\subsection{Hypersoundness}\label{hypersoundeness}

Now that we proved that regular POP converges, we can introduce the next
property: hyper soundness.

\begin{definition}[Hypersoundness]

An algorithm is said to be hypersound when it gives a valid solution for
all problems regardless of their resolvability.

\end{definition}

We note that this property isn't compatible with consistency regarding
the original problem and then doesn't fit the classical idea of
soundness that implicitly states that the validity of a solution is
relative to the problem. In the case of hypersoundness, we find a
solution to a derived problem \(P' = \langle A', I, G, p\rangle\). We
note the new set \(A' = A \cup S\) with \(S\) being the set of saviors
for all flaws that made the POP algorithm fail during the execution of
SODA. Next we prove the hypersoundness of our algorithm using the
convergence of POP and the way the Soft solving behaves.

\begin{proof}[Proof of hypersoundness of SODA POP]

The proper plan and defect fixing algorithms are obviously convergent.
The proper plan algorithm cannot iterate more than the number of actions
(since duplicates are forbidden) and the defect resolution will fix the
finite number of defects present in the finite partial plan issued by
the proper plan algorithm.

POP is sound and converges. Therefore, if SODA POP converges it will
return a valid solution for \(P'\). In the same way we proved for POP,
the new healing process converges because it reduces the number of flaws
in the partial plan and since this number is finite, the algorithm
converges.\qedhere

\end{proof}

\subsection{Enhancement for online
planning}\label{enhancement-for-online-planning}

SODA POP algorithm can be used in dynamic online environments to allow a
robust way to replan an existing obsolete plan. The first step prior to
the runtime is the generation of the proper plans for all the goals that
will be considered. For each online iteration, we take the previous
plan, clean all defects in it and makes it loop between POP and the
healing mechanism (if necessary). The defect cleaning will also remove
all unnecessary savior actions and eventually make the original problem
solvable again. Since each previous plan is almost correct, the
algorithm will have little to no iterations to make in order to repair
it.

\section{Experimental results}\label{experimental-results}

In order to verify the validity of these algorithms we implemented a
prototype in Java. We used basic actions with integer fluents to focus
on the way the algorithm behaves. For the simulation runs, we used a
computer with an Intel® Core™ i7-4720HQ CPU clocking 8 cores at 2.60GHz
and 8GB of memory. The speed was measured in nano seconds before and
after the call to the solve function. The process has a warming up phase
that computes random plans for a few seconds before starting the
benchmarks.

\subsection{Quality of classical POP}\label{quality-of-classical-pop}

The first metric we are interested in is the performance of classical
POP algorithm. We measure the solving time for valid problems and the
quality of the resulting plans. We tested the algorithm with randomly
generated valid problems. We generated those by randomly creating plans
based on a difficulty setting that was the upper bound for the number of
actions and half the number of possible fluents. This difficulty setting
also made the initial and goal step larger accordingly. The actions were
cleaned of any toxicity and were built using a forward chaining
algorithm. We added also random unrelated actions. Each problem was
verified by POP before being generated to ensure resolvability. As we
clearly see in the figure \ref{fig:quality}, POP linearly increases its
solving time depending on the problem difficulty. That obviously follows
the fact that POP has more flaws to fix and therefore more iterations.
As we can observe, POP has a rather high link quality with almost no
problems issued with causal link regardless of the difficulty of the
problem. This can be linked to either the way we generated our valid
plans or to a real tendency of POP to not create that kind of defects on
its own. The action quality on the other hand drops significantly as the
difficulty gets higher. This is linked to the number of competing and
useless actions that can make the plan simpler but are kept by the flaw
selection mechanism of POP.

\begin{figure}[htbp]
\centering
\includegraphics{graphics/quality.pdf}
\caption{Action and link quality and solving time of POP over \(10^4\)
valid problems for each difficulty\label{fig:quality}}
\end{figure}

\subsection{Performance of SODA POP}\label{performance-of-soda-pop}

As SODA POP and regular POP have not the same range of capabilities we
can't compare them hand to hand. Indeed, our algorithm always outputs
plans with 100\% quality since the defect detection system aims to
remove all of them. Moreover, our algorithm is hypersound so it gives
valid plans for unsolvable problems by derivation. In this specific case
POP won't be able to return a result and terminates more quickly as the
plan is found unsolvable.

Therefore, we measured the performances of SODA POP algorithm on
completely random problems. Each time the algorithm outputs a solution
with the lowest possible violation despite the complete invalidity of
the input. In figure \ref{fig:performance} we can see the way SODA POP
scales up on larger problems. The performances display logarithmic
solving time. This is explained by the fact that POP has to be called
recursively on the healed problems. The defect detection participates
too in the increase of solving time especially the competing action
detection that needs to iterate over all actions and edges and calculate
usefulness for two actions each time.

\begin{figure}[htbp]
\centering
\includegraphics{graphics/performance.pdf}
\caption{Average solving time of SODA POP over \(10^4\) random problems
for each difficulty\label{fig:performance}}
\end{figure}

\section*{Conclusion}\label{conclusion}
\addcontentsline{toc}{section}{Conclusion}

We defined our SODA POP algorithm and demonstrated its properties. The
set of algorithms has extended capabilities of plan repairing and soft
solving as shown in examples and simulations. While slower than POP, the
resilience of SODA POP can be used in various applications. We aim to
use it on the particular case of online plan recognition and decision
making as the capabilities to derive plans and repair them will prove
useful when comparing potential goals with observed plans. The algorithm
can be improved with a better defect detection algorithm that can rank
defects and start fixing the most offensive defects before targeting
others. Also, if the competing action detection can be improved, it can
be used as heuristic for resolver selection to improve the performances.
Another way the present work can be improved is by extending it to the
use of more expressive fluents and transpose all notions to variable
domains. One possible goal would be to improve fluent domain with an
propositional language in order to significantly boost expressiveness.
Finally, we plan to release the software with an open source license
soon.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-weldux5fintroductionux5f1994}{}
{[}1{]} D. S. Weld, ``An introduction to least commitment planning,''
\emph{AI magazine}, vol. 15, no. 4, p. 27, 1994.

\hypertarget{ref-richterux5flamaux5f2011}{}
{[}2{]} S. Richter, M. Westphal, and M. Helmert, ``LAMA 2008 and 2011,''
in \emph{International Planning Competition}, 2011, pp. 117--124.

\hypertarget{ref-ghallabux5fautomatedux5f2004}{}
{[}3{]} M. Ghallab, D. S. Nau, and P. Traverso, \emph{Automated
planning: theory and practice}. Amsterdam ; Boston: Elsevier/Morgan
Kaufmann, 2004.

\hypertarget{ref-vanux5fderux5fkrogtux5fplanux5f2005}{}
{[}4{]} R. Van Der Krogt and M. De Weerdt, ``Plan Repair as an Extension
of Planning.'' in \emph{ICAPS}, 2005, vol. 5, pp. 161--170.

\hypertarget{ref-ramirezux5fplanux5f2009}{}
{[}5{]} M. Ramırez and H. Geffner, ``Plan recognition as planning,'' in
\emph{Proceedings of the 21st international joint conference on
Artifical intelligence. Morgan Kaufmann Publishers Inc}, 2009, pp.
1778--1783.

\hypertarget{ref-erolux5fumcp:ux5f1994}{}
{[}6{]} K. Erol, J. A. Hendler, and D. S. Nau, ``UMCP: A Sound and
Complete Procedure for Hierarchical Task-network Planning.'' in
\emph{AIPS}, 1994, vol. 94, pp. 249--254.

\hypertarget{ref-penberthyux5fucpop:ux5f1992}{}
{[}7{]} J. S. Penberthy, D. S. Weld, and others, ``UCPOP: A Sound,
Complete, Partial Order Planner for ADL.'' \emph{Kr}, vol. 92, pp.
103--114, 1992.

\hypertarget{ref-bechonux5fhipop:ux5f2014}{}
{[}8{]} P. Bechon, M. Barbier, G. Infantes, C. Lesire, and V. Vidal,
``HiPOP: Hierarchical Partial-Order Planning,'' in \emph{STAIRS 2014:
Proceedings of the 7th European Starting AI Researcher Symposium}, 2014,
vol. 264, p. 51.

\hypertarget{ref-garciaux5fdefeasibleux5f2008}{}
{[}9{]} D. R. García, A. J. García, and G. R. Simari, ``Defeasible
reasoning and partial order planning,'' in \emph{Foundations of
Information and Knowledge Systems}, Springer, 2008, pp. 311--328.

\hypertarget{ref-colesux5fforward-chainingux5f2010}{}
{[}10{]} A. J. Coles, A. Coles, M. Fox, and D. Long, ``Forward-Chaining
Partial-Order Planning.'' in \emph{ICAPS}, 2010, pp. 42--49.

\end{document}